Lab2

### 2021-08-04 src/data/DataCleaning.Rmd

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim.csv')
```

```{r merged model_1 dataframe, echo=FALSE}
#code to merge county info to d
d <- merge(x = d_trips, y = d_covid, by = c("fips", "county", "state"))

#drop na and duplicates
d <- d %>% drop_na()
d <- unique(d)
```

@@

@@

# Clean other variables
```{r read in variable data, echo=FALSE}
density <- read.delim("./../../data/raw/2019_Gaz_counties_national.txt", header=TRUE, sep="\t")
dp03_econ <- read.csv(file = './../../data/raw/ACSDP5Y2019.DP03_data_with_overlays_2021-07-26T143523.csv')
dp05_age <- read.csv(file = './../../data/raw/ACSDP1Y2019.DP05_data_with_overlays_2021-07-21T215013.csv')
s1901_income <- read.csv(file = './../../data/raw/ACSST1Y2019.S1901_data_with_overlays_2021-07-21T222906.csv')

unemp_claims <- readxl::read_excel("./../../data/raw/weekly-claims-by-county-twc.xlsx")
ur <- readxl::read_excel("./../../data/raw/TX_Rural_County.xlsx")
```

```{r clean dp03 data, echo=FALSE}
econ <- dp03_econ %>%
  rename(fips = GEO_ID, unemp_rate = DP03_0009PE) %>%
  select(fips, unemp_rate) %>%
  mutate(
    fips = substring(fips, 10),
    unemp_rate = as.double(unemp_rate)
  )

# merge into main dataframe
d <- merge(x = econ, y = d, by = c("fips"))
```

```{r clean twc unemployment data, echo=FALSE}

uclaims <- unemp_claims[-c(1,2),] %>%
  select(TEXAS,
         ...6,
         ...7,
         ...8,
         ...9
         ) %>% 
  rowwise() %>%
  mutate(
    ...6 = as.integer(...6),
    ...7 = as.integer(...7),
    ...8 = as.integer(...8),
    ...9 = as.integer(...9),
    total_claims = sum(c(...6, ...7, ...8, ...9))
  ) %>%
  rename(
    county = TEXAS
  ) %>% drop_na()

# merge into main dataframe
d <- merge(x = select(uclaims, c(county, total_claims)), y = d, by = c("county"))
```

```{r clean dp03 data, echo=FALSE}
# filters by date/state
age <- dp05_age %>%
  rename(fips = GEO_ID,
         female = DP05_0003PE, 
         pct_lt5 = DP05_0005PE, 
         pct_5to9 = DP05_0006PE, 
         pct_10to14 = DP05_0007PE, 
         pct_15to19 = DP05_0008PE, 
         pct_20to24 = DP05_0009PE, 
         pct_25to34 = DP05_0010PE, 
         pct_35to44 = DP05_0011PE, 
         pct_45to54 = DP05_0012PE
) %>% rowwise() %>%
  mutate(
    fips = substring(fips, 10),
    female = as.double(female), 
    pct_lt5 = as.double(pct_lt5), 
    pct_5to9 = as.double(pct_5to9), 
    pct_10to14 = as.double(pct_10to14), 
    pct_15to19 = as.double(pct_15to19), 
    pct_20to24 = as.double(pct_20to24), 
    pct_25to34 = as.double(pct_25to34), 
    pct_35to44 = as.double(pct_35to44), 
    pct_45to54 = as.double(pct_45to54),
    pct_gt55 = 100 - sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19, pct_20to24, pct_25to34, pct_35to44, pct_45to54)),
    pct_lt19 = sum(c(pct_lt5, pct_5to9, pct_10to14, pct_15to19)),
    pct_19to54 = 100 - pct_gt55 - pct_lt19
    ) %>%
  select(fips, pct_lt19, pct_19to54, pct_gt55)

# merge into main dataframe
d <- merge(x = age, y = d, by = c("fips"))
```

```{r clean s1901 data, echo=FALSE}
# filters by date/state
income <- s1901_income %>%
  rename(fips = GEO_ID,
  median_income = S1901_C01_012E,  
  pct_50to75k = S1901_C01_007E, 
  pct_75to100k = S1901_C01_008E, 
  pct_100to150k = S1901_C01_009E, 
  pct_150to200k = S1901_C01_010E, 
  pct_gt200k = S1901_C01_011E
) %>% rowwise() %>%
  mutate(
    fips = substring(fips, 10),
    median_income = as.double(median_income), 
    pct_50to75k = as.double(pct_50to75k), 
    pct_75to100k = as.double(pct_75to100k), 
    pct_100to150k = as.double(pct_100to150k), 
    pct_150to200k = as.double(pct_150to200k), 
    pct_over_us_median = sum(c(pct_50to75k, pct_75to100k, pct_100to150k, pct_150to200k))) %>%
  select(fips, median_income, pct_over_us_median)

# merge into main dataframe
d <- merge(x = income, y = d, by = c("fips"))
```

```{r rural indicator, echo=FALSE}
d$rural_ind <- ifelse(d$county %in% ur$`Rural County`, 1, 0)

summary(d$rural_ind)

@@

@@

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim.csv')
```

>>>>

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_limited.csv')
```

@@

```{r load unemployment data}
unemploy <- read.csv('./../../data/raw/TAC_unemployment.csv')
```

```{r clean unemployment data}
unemploy_f <- unemploy %>%
  rename(
    county = County,
    unempoy_rate = Unemployment.Rate....
  ) %>%
  select(
    county, 
    unempoy_rate 
  )

```{r merge model 2 dataframe}
d <- merge(x = unemploy_f, y = d, by = c("county"))
d <- d %>% drop_na()

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_model2.csv')
```

@@

```{r load age data}
age <- read.csv('./../../data/raw/TAC_age.csv')
```

```{r clean unemployment data}
age_f <- age %>%
  rename(
    county = County,
    pct_lt17 = X..Age.17.and.Under
  ) %>% rowwise() %>%
  mutate(
    pct_gt65 = sum(c(X..Age.65.and.Older, X..Age.85.and.Older))
  ) %>%
  select(
    county,
    pct_lt17,
    pct_gt65
  )

@@

```{r merge limited model dataframe}
d <- merge(x = age_f, y = d, by = c("county"))
d <- d %>% drop_na()

# Write to .csv
```{r export interim model_1 dataframe as csv, echo=FALSE}
write.csv(d, '../../data/interim/covid_trips_interim_model3.csv')
```

@@
